Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece

### 2023-11-09 Build_163/MDPI_submission_B163.Rmd
Commit 2023-11-09 19:08

---
title:     "Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece"
author:
  - name:  Athanasios Natsis
    affil: 1,* # \ddagger
    orcid: 0000-0002-5199-4119
  - name: Alkiviadis Bais
    affil: 1 #, \dagger, # \ddagger
  - name: Alkiviadis Bais
    affil: 1 #, \dagger, # \ddagger

affiliation:
  - num: 1
    address: |
      Aristotle University of Thessaloniki -
      Laboratory of Atmospheric Physics,
      Campus Box 149, 54124 Thessaloniki, Greece
    email: natsisphysicist@gmail.com (A.N.); abais@auth.gr (A.B.)
  # - num: 2
  #   address: |
  #     Your department
  #     Street, City, Country
#     email: abais@auth.gr
# author citation list in chicago format
authorcitation: |
  Natsis, A.; Bais, A.
# firstnote to eighthnote
# firstnote: |
#   Current address: Updated affiliation
# secondnote: |
#   These authors contributed equally to this work.
correspondence: |
  natsisphysicist@gmail.com;
#   Tel.: +XX-000-00-0000.

@@

abstract: |
  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
  Observations of SDR at different locations around the world with different environmental characteristics have been used to investigate its long-term variability and trends at different time scales.
  Periods of positive trends are referred as brightening periods and of negative trends as dimming periods.
  Here we studied 30 years of pyranometer data in Thessaloniki, Greece, under three types of sky conditions (clear sky, cloudy sky and all sky).
  The clear-sky data were identified by applying a cloud screening algorithm. We found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $0.35\,\%/\text{year}$ for clear-sky conditions, and $-0.28\,\%/\text{year}$ for cloudy conditions.
  We have also investigated the consistency of these trends, their seasonal variability, and the effect of the solar zenith angle. 
  We have found that for all-sky and clear-sky conditions the SDR trend is positive in winter (0.7 and 0.8 %/y, respectively) and autumn (~0.4 %/y), while under cloudy skies the trend is negative (-0.9 %/y in winter and -0.4 %/y in autumn). 
  In spring and summer the trend is very close to zero, irrespective of sky conditions. 
  The SDR trend is increasing with increasing solar zenith angle, except under cloudy skies where the trend is highly variable and close to zero.
  Finally, we identified some anomalies in the long term SDR trends for all sky conditions by examining the cumulative sums of monthly anomalies from the climatological mean.

>>>>

---
title:     "Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece"
author:
  - name:  Athanasios Natsis
    affil: 1,* # \ddagger
    orcid: 0000-0002-5199-4119
  - name: Alkiviadis Bais
    affil: 1 #, \dagger, # \ddagger
  - name: Charikleia Meleti
    affil: 1 #, \dagger, # \ddagger

affiliation:
  - num: 1
    address: |
      Aristotle University of Thessaloniki -
      Laboratory of Atmospheric Physics,
      Campus Box 149, 54124 Thessaloniki, Greece
#    email: natsisphysicist@gmail.com (A.N.); abais@auth.gr (A.B.); meleti@auth.gr (C.M.)
  # - num: 2
  #   address: |
  #     Your department
  #     Street, City, Country
#     email: abais@auth.gr
# author citation list in chicago format
authorcitation: |
  Natsis, A.; Bais, A.; Meleti, C.
# firstnote to eighthnote
# firstnote: |
#   Current address: Updated affiliation
# secondnote: |
#   These authors contributed equally to this work.
correspondence: |
  natsisphysicist@gmail.com;
#   Tel.: +XX-000-00-0000.

@@

abstract: |
  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
  Observations of SDR at different locations around the world with different environmental characteristics have been used to investigate its long-term variability and trends at different time scales.
  Periods of positive trends are referred as brightening periods and of negative trends as dimming periods.
  Here we studied 30 years of pyranometer data in Thessaloniki, Greece, under three types of sky conditions (clear sky, cloudy sky and all sky).
  The clear-sky data were identified by applying a cloud screening algorithm. We found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $0.35\,\%/\text{year}$ for clear-sky conditions, and $-0.28\,\%/\text{year}$ for cloudy conditions.
  We have also investigated the consistency of these trends, their seasonal variability, and the effect of the solar zenith angle. 
  We have found that for all-sky and clear-sky conditions the SDR trend is positive in winter ($0.7$ and $0.8\,\%/\text{year}$, respectively) and autumn ($~0.4\,\%/\text{year}$), while under cloudy skies the trend is negative ($-0.9\,\%/\text{year}$ in winter and $-0.4\,\%/\text{year}$ in autumn). 
  In spring and summer the trend is very close to zero, irrespective of sky conditions. 
  The SDR trend is increasing with increasing solar zenith angle, except under cloudy skies where the trend is highly variable and close to zero.
  Finally, we identified some anomalies in the long term SDR trends for all sky conditions by examining the cumulative sums of monthly anomalies from the climatological mean.

### 2023-11-09 Build_163/MDPI_submission_B163.tex
Commit 2023-11-09 19:08

% Authors, for the paper (add full first names)
\Author{Athanasios
Natsis$^{1,*}$\href{https://orcid.org/0000-0002-5199-4119}
{\orcidicon}, Alkiviadis Bais$^{1}$}

@@

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Athanasios Natsis, Alkiviadis Bais}

% MDPI internal command: Authors, for citation in the left column
%\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
\AuthorCitation{Natsis, A.; Bais, A.}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Aristotle University of Thessaloniki - Laboratory of
Atmospheric Physics, Campus Box 149, 54124 Thessaloniki,
Greece; \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}
(A.N.); \href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}} (A.B.)\\
}

@@

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{The shortwave downward solar irradiance (SDR) is an important
factor that drives climate processes, production and can affect all
living organisms. Observations of SDR at different locations around the
world with different environmental characteristics have been used to
investigate its long-term variability and trends at different time
scales. Periods of positive trends are referred as brightening periods
and of negative trends as dimming periods. Here we studied 30 years of
pyranometer data in Thessaloniki, Greece, under three types of sky
conditions (clear sky, cloudy sky and all sky). The clear-sky data were
identified by applying a cloud screening algorithm. We found a positive
trend of \(0.38\,\%/\text{year}\) for all-sky, \(0.35\,\%/\text{year}\)
for clear-sky conditions, and \(-0.28\,\%/\text{year}\) for cloudy
conditions. We have also investigated the consistency of these trends,
the effect of the solar zenith angle, and their seasonal variability.
Finally, we identified some anomalies in the long term SDR trends for
all sky conditions by examining the cumulative sums of monthly anomalies
from the climatological mean.}

>>>>

% Authors, for the paper (add full first names)
\Author{Athanasios
Natsis$^{1,*}$\href{https://orcid.org/0000-0002-5199-4119}
{\orcidicon}, Alkiviadis Bais$^{1}$, Charikleia Meleti$^{1}$}

@@

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Athanasios Natsis, Alkiviadis Bais, Charikleia Meleti}

% MDPI internal command: Authors, for citation in the left column
%\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
\AuthorCitation{Natsis, A.; Bais, A.; Meleti, C.}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Aristotle University of Thessaloniki - Laboratory of
Atmospheric Physics, Campus Box 149, 54124 Thessaloniki, Greece; \\
}

@@

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{The shortwave downward solar irradiance (SDR) is an important
factor that drives climate processes, production and can affect all
living organisms. Observations of SDR at different locations around the
world with different environmental characteristics have been used to
investigate its long-term variability and trends at different time
scales. Periods of positive trends are referred as brightening periods
and of negative trends as dimming periods. Here we studied 30 years of
pyranometer data in Thessaloniki, Greece, under three types of sky
conditions (clear sky, cloudy sky and all sky). The clear-sky data were
identified by applying a cloud screening algorithm. We found a positive
trend of \(0.38\,\%/\text{year}\) for all-sky, \(0.35\,\%/\text{year}\)
for clear-sky conditions, and \(-0.28\,\%/\text{year}\) for cloudy
conditions. We have also investigated the consistency of these trends,
their seasonal variability, and the effect of the solar zenith angle. We
have found that for all-sky and clear-sky conditions the SDR trend is
positive in winter (\(0.7\) and \(0.8\,\%/\text{year}\), respectively)
and autumn (\(~0.4\,\%/\text{year}\)), while under cloudy skies the
trend is negative (\(-0.9\,\%/\text{year}\) in winter and
\(-0.4\,\%/\text{year}\) in autumn). In spring and summer the trend is
very close to zero, irrespective of sky conditions. The SDR trend is
increasing with increasing solar zenith angle, except under cloudy skies
where the trend is highly variable and close to zero. Finally, we
identified some anomalies in the long term SDR trends for all sky
conditions by examining the cumulative sums of monthly anomalies from
the climatological mean.}

### 2023-11-09 SUBMISSION/appendix.tex
Commit 2023-11-09 07:20
Commit 2023-11-09 19:08

%% optional
\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixsections{one} %Leave argument "multiple" if there are multiple sections. Then a counter is printed ("Appendix A"). If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
\appendix
\section{}

% \subsection{}

\  \

% copy output from Article.tex

@@

\  \

@@

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        \centering
        \includegraphics[width=1.0\linewidth]{./images/SzaTrendsSeasTogether-2}
        \caption{Long term trends of SDR as a function of SZA separately form morning
               and afternoon periods, by season (rows of plots) for the three sky
               conditions (columns of plots).
               Solid shapes represent statistical significant trends ($p<0.005$).
               Cases where $p<0.005$ or with less than $85$ observations may be
               missing from view.}\label{fig:SZAtrendSeason}
    \end{adjustwidth}
\end{figure}

% \begin{figure}
%     \begin{adjustwidth}{-\extralength}{0cm}
%         \centering
%         \includegraphics[width=1.0\fulllength]{./images/SzaTrendsSeasTogether-2}
%     \end{adjustwidth}
%     \caption{This figure will be full-page.\label{fig:panoramic}}
% \end{figure}
%
% \begin{figure}
%     \begin{adjustwidth}{-\extralength}{0cm}
%         \centering
%         \includegraphics[width=1.0\linewidth]{./images/SzaTrendsSeasTogether-2}
%     \end{adjustwidth}
%     \caption{This figure will be full-page.\label{fig:panodramic}}
% \end{figure}

>>>>

% % \renewcommand\thefigure{\thesection.\arabic{figure}} %% my
% % \renewcommand\thefigure{\thesection.\arabic{figure}}
%
% %% From R rticle MDPI template
% %% optional
% \appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
% \appendixsections{one} %Leave argument "multiple" if there are multiple sections. Then a counter is printed ("Appendix A"). If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
%
% % \appendix
% % \section{}
%
% \appendix
% \setcounter{secnumdepth}{0}
% \section{Appendix}
%
% \setcounter{figure}{0}    %% my

%% From raw MDPI template
%% Optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixstart
\appendix
\section[\appendixname~\thesection]{}

% copy output from Rmd

@@

@@

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering
            \includegraphics[width=1.0\linewidth]{./images/SzaTrendsSeasTogether-2}
        }
        \caption{Long term trends of SDR as a function of SZA separately form morning
               and afternoon periods, by season (rows of plots) for the three sky
               conditions (columns of plots).
               Solid shapes represent statistical significant trends ($p<0.005$).
               Cases where $p<0.005$ or with less than $85$ observations may be
               missing from view.}\label{fig:SZAtrendSeason}
    \end{adjustwidth}
\end{figure}
